
\section{Alternative Numeraires}

So far we have discounted the price process by a riskless bank account.
\[A_t^*=\frac{A_t}{A_t}\]
\[G_t^*=S_t^*+D_t^*=\frac{S_t}{A_t}^*+\int_t^T\frac{\delta_S(u)}{A_u}du\]
We then chose $Q$ such that $G_t^*$ was a $Q$ martingale, $G_t^*=E_t^Q[G_T^*]$.
We don't have to use a bank account as the numeraire.
Depending on the goals, other securities may save some computation time and are
therefore better.

The requirements for a general numeraire security are; strictly positive and no
separate dividends, i.e. $B$ is a gains process for $B$.
\[A_t^B=\frac{A_t}{B_t}\]
\[G_t^B=S_t^B+D_t^B=\frac{S_t}{B_t}+\int_t^T\frac{\delta_S(u)}{B_u}du\]
We then chose $Q_B$ such that $G_t^B$ was a $Q$ martingale,
$G_t^B=E_t^{Q_B}[G_T^B]$.

\section{Forward Measure, $Q_T$}

$P(t,T)$ is the date $t$ price of a ZCB paying $1$ at time $T$. We define the
$T$ forward measure, $Q_T$, by requiring
$G_t^{P(t,T)}=E_t^{Q_T}[G_s^{P(t,T)}]$ for all $t\leq s$ for discounted gains
processes. That is to say, $Q_T$ is the probability measure that makes all
gains processes martingales.

\subsection{Example: EU Call}

\[c_t^{P(t,T)}=E_t^{Q_T}[c_T^{P(t,T)}]\]
iff
\[\frac{c_t}{P(t,T)}=E_t^{Q_T}\left[\frac{max(S_T-K,0)}{P(T,T)}\right]\]
$P(T,T)$ is $1$ by construction and $P(t,T)$ can be observed
\[c_t=E_t^{Q_T}[max(S_T-K,0)]P(t,T)\]
Who cares? This is the discounted expected value as before. But the forward
measure helps if interest rates are stochastic. Under the spot measure we would
have the expectation of a random variable times a random variable. This is
a covariance nightmare.

\section{Stock Measure, $Q_S$}

While the stock price process from our model is strictly positive, it does pay
dividends. We must adjust it slightly to use it as a numeraire. We use an
artificial security $S^c$ where all dividends are continuously reinvested. The
intuition is similar to the bank account. We accumulate units a the rate of
dividend payout (like we accumulate dollars at the risk free rate).
\begin{equation} \label{stockNumeraire}
  S_te^{\int_0^t\delta_S(u)du}
\end{equation}
Who cares? In short, $G^S$ will be $Q$-martingales by construction. More
thoroughly, We know our stock price process $S$ has a gains process of
$G=S+D$ and we know $S_t=S_0e^{\left(\mu-\frac{1}{2}\sigma^2\right) t+
\sigma W_t}$, $dS_t=\mu S_tdt+\sigma S_tdW_t$ and $dD_t=\delta S_tdt$.
Dividing through by $S_t$ removes the $\mu$. But the price process is still
growing at the dividend rate, $\delta$. Dividing by the adjusted stock price
process, $S^c$, removes $\mu$ and $\delta$ making $G^S$ a $Q$-martingale by
construction. 
While we are using the stock measure we can still use the bank account to find
the correct risk adjustment, $\lambda^S$.
\[
  A_t^S=\frac{A_t}{S_t^c}=\frac{1}{S_t}
  \frac{e^{\int_0^tr_udu}}{e^{\int_0^t\delta_S(u)du}}
\]
\[=\frac{1}{S_t}e^{\int_0^tr_u-\delta_S(u)du}\]
Now we need to find the dynamics of this diffusion, It\^o to the rescue.
Identification:
\[f(t,S)=\frac{1}{S_t}e^{\int_0^tr_u-\delta_S(u)du}\]
The $S$ means we need It\^o because this is a diffusion. Calculation:
\[f_1=\frac{1}{S_t}e^{\int_0^tr_u-\delta_S(u)du}(r_t-\delta_S(t))\]
\[f_2=-\frac{1}{S_t^2}e^{\int_0^tr_u-\delta_S(u)du}\]
\[f_{22}=2\frac{1}{S_t^3}e^{\int_0^tr_u-\delta_S(u)du}\]
Substitution (into \ref{IL}):
\[dA_t^S=\]
\[
  \left[ \frac{1}{S_t}e^{\int_0^tr_u-\delta_S(u)du}(r_t-\delta_S(t))-
  \frac{1}{S_t^2}e^{\int_0^tr_u-\delta_S(u)du}\mu_S(t)S_t+
  \frac{1}{2}2\frac{1}{S_t^3}e^{\int_0^tr_u-\delta_S(u)du}(\sigma_S(t)S_t)^2
  \right] dt+
\]
\[-\frac{1}{S_t^2}e^{\int_0^tr_u-\delta_S(u)du}\sigma_S(t)S_tdW_t\]
Simplification:
\[
  dA_t^S=\left[ \frac{A_t}{S_t^c}(r_t-\delta_S(t))-\frac{A_t}{S_t^c}\mu_S(t)+
  \frac{A_t}{S_t^c}\sigma_S^2(t) \right] dt-\frac{A_t}{S_t^c}\sigma_S(t)dW_t
\]
Adjust the Wiener process:
\[
  dA_t^S=A_t^S\left[ r_t-\delta_S(t)-\mu_S(t)+\sigma_S^2(t)\right] dt-
  A_t^S\sigma_S(t)\left( dW_t^S-\lambda_t^Sdt\right)
\]
\[
  dA_t^S=A_t^S\left[ r_t-\delta_S(t)-\mu_S(t)+\sigma_S^2(t)+
  \sigma_S(t)\lambda_t^S\right] dt-A_t^S\sigma_S(t)dW_t^S
\]
is a $Q_S$-martingale iff no drift, i.e. set $dt$ term to $0$ and solve for
$\lambda$.
\[\lambda_t^S=\frac{-r_t+\delta_S(t)+\mu_S(t)-\sigma_S^2(t)}{\sigma_S(t)}\]
Now that we have nailed down $\lambda$ in the gains process we can use it to
adjust the stock price process from $P$ to $Q_S$ probabilities.
\[dS_t=\mu_S(t)S_tdt+\sigma_S(t)S_tdW_t\]
\[dS_t=\mu_S(t)S_tdt+\sigma_S(t)S_t(dW_t^S-\lambda_t^Sdt)\]
\[
  dS_t=\mu_S(t)S_tdt+\sigma_S(t)S_t\left[ dW_t^S-
  \frac{-r_t+\delta_S(t)+\mu_S(t)-\sigma_S^2(t)}{\sigma_S(t)}\right]
\]
\[dS_t(r_t-\delta_S(t)+\sigma_S^2(t))S_tdt+\sigma_S(t)S_tdW_t^S\]
We recognise this as a geometric Wiener.
