
\section{Systemic Risk in Option Pricing}

Finding $\lambda$ is independent of derivative payoff we are valuing. The
underlying asset has systemic risk embodied by $\mu_S$ and $\delta_S$. They do
not show up in the valuation directly. It is not immediately intuitive why we
can price different derivatives, with different exposures to systemic risk,
using the same $\lambda$. Do all derivatives on $S$ have the same systemic
risk? \textbf{No}.

To understand how we can use $\lambda$ indifferently we need to look at the
relationship between $P$ and $Q$ measures.
$f_P$ and $f_Q$ are the probability densities associated with $P$ and $Q$,
respectively. We can say that $f_Q=zf_P$. Where
$$z=\frac{\partial Q}{\partial P}=
e^{-\frac{1}{2}\int_0^t\lambda_u^2du-\int_0^t\lambda_udW_u}$$
$z$ is a function of $\lambda$ and therefore of the Sharpe ratio under each
probability measure.
\[c_t=E_t^Q[c_T]\]
\[=\int_{-\infty}^{\infty}c_T(\varepsilon)f_Q(\varepsilon)d\varepsilon\]
\[
  =\int_{-\infty}^{\infty}c_T(\varepsilon)z(\varepsilon)
  f_P(\varepsilon)d\varepsilon
\]
\[=E_t^P[c_Tz]\]
But a product inside an expectation means there is covariance lurking.
\[=E_t^P[c_T]E_t^P[z]+Cov_t^P(c_T,z)\]
where $E_t^P[z]=1$
\[=E_t^P[c_T]+Cov_t^P(c_T,z)\]
The covariance is punishment and project specific. The $z$ may be the same
between projects but the cashflows can differ. No arbitrage accounts for the
$z$.

When we construct an EMM we are dragging the drift of a diffusion to $0$. Under
$P$ probabilities the diffusion is described by
\[S_t=S_0e^{(\mu-\frac{1}{2}\sigma^2)t-\sigma W_t}\]
and the expected value is
\[E_0^P[S_t]=S_0e^{\mu t}\]
When we change to $Q$ probabilities we drag the growth rate down
\[S_t=S_0e^{(r-\frac{1}{2}\sigma^2)t-\sigma\tilde{W}_t}\]
and the expected value drops to
\[E_0^Q[S_t]=S_0e^{rt}\]
We make the diffusion a martingale by dividing by the numeraire security
\[A_t=e^{rt}\]
\[E_0^Q\left[\frac{S_t}{A_t}\right]\]
$r$ is compensation for impatience while $\lambda\sigma$ is compensation for
risk.
