
\section{Generalized Weiner Process}

The standard Wiener process is just a series of numbers drawn from a
distribution (white noise from ECO403). We can generalize this somewhat by
adding a deterministic drift component and a deterministic uncertainty scaling
to the source of uncertainty provided by $W$. In differential form this looks
like:
\begin{equation} \label{dGenW}
    dX_t=\mu_X(t)dt+\sigma_X(t)dW_t
\end{equation}
where $\mu_X(t)dt$ is the drift, $\sigma_X(t)$ is the uncertainty scaling and
$dW_t$ is the source of uncertainty (our increments of W from above
$\sim N(0, t-s)$).
In descrete time this looks like:
\begin{equation}
    X_{t+\Delta t}-X_t=\Delta X_t=\mu_X(t)\Delta t+\sigma_X(t)\Delta W_t
\end{equation}
Since $E[\Delta W_t]=0$, Expected value of $\Delta X_t$ is
\begin{equation} \label{ExpDesGenW}
    E[\Delta X_t]=\mu_X(t)\Delta t
\end{equation}
Since variance of a constant is 0 and the variance of the standard Wiener
process, W, is $t-s=\Delta t$ AND the two terms are independent of each other,
we can ignore the covariance (and remembering that constants taken outside of
$Var(...)$ must be squared)
\begin{equation} \label{VarDesGenW}
    Var(\Delta X_t)=\sigma^2_X(t)\Delta t
\end{equation}
Linear combinations of normal random variables are normal. Therefore,
\begin{equation}
    \Delta X_t \sim N(\mu_X(t)\Delta t, \sigma^2_X(t)\Delta t)
\end{equation}
\begin{equation}
    dX_t \sim N(\mu_X(t)dt, \sigma^2_X(t)dt)
\end{equation}

If we wanted to solve X we could devide through by dt. Problem is
$\frac{dW_t}{dt}$ does not exist because otherwise we could trade based on the
derivative and then there would be arbitrage and wailing and gnashing of teeth.
We can write $dX_t$ in integral form
\begin{equation} \label{GenW}
    X_t=x+\int_0^t\mu_X(u)du+\int_0^t\sigma_X(u)dW_u
\end{equation}
The funny $dW_u$ means this is an It\^o integral. In fancy language, ``The time
period is transformed by the stochastic process." I do not know what this
means. Basically, the point here is that we cannot use regular integration to
find $X_t$. But, someone clever called It\^o has got our backs here as we'll
see later. Nevertheless, we can say something about the expected value and
variance of $X_t$ using descrete time intuition as we did for $dX_t$.

The expected value of $X_t$ is the expected value of $x$, which we just set
ourselves, plus all the drift terms multiplied by the time interval. We can
ignore the dispersion term because $W_t$ comes from a distribution with mean 0.
We are summing up the expectation of $\Delta X_t$ we came up with in equation
\ref{ExpDesGenW}.
\begin{equation}
    E[X_t]=E[x]+E \left[ \sum \mu_X(u)\Delta u\right] +0
\end{equation}

The deterministic terms have no variance so we can ignore them. The increments
of $W$ are independent so we do not need to think about covariances here. The
variance of a sum is the sum of the variances. We can sum up the descrete time
variances of $\Delta X_t$ we came up with above (remembering the $\sigma_X^2$
was introduced when the deterministic uncertainty scaling component
$\sigma_X(t)$ was taken outside of the $Var(...)$ term and so had to be
squared in equation \ref{VarDesGenW}. $\Delta t=t-s$, $\Delta u$ takes the
place of $\Delta t$ here.)
\begin{equation}
    Var(X_t)=\sum \sigma^2(u)\Delta u
\end{equation}

Linear combinations of normal random numbers are normal. Putting this back
into continuous time gives us:
\begin{equation}
    X_t \sim N \left( x+ \int_0^t \mu_X(u)du, \int_0^t \sigma_X^2(u)du \right)
\end{equation}

X has inherited; independent increments and continuous, non-differentiable
sample paths. We have not finished. X can still be negative which is not
consistent with limited liability of stocks. Furthermore, we still cannot solve
the funny It\^o integral $\frac{dW_u}{du}$. We need to use a process for which
a solution to the It\^o integral has already been found that has the properties
that we have just described above; drift, uncertainty scaling, a source of
uncertainty and independent increments.
