
\section{Spot Measure Black-Scholes}

Using our spot measure where the numeraire asset is the bank account gave us;
\[dA_t=r_tA_tdt\]
\[dS_t=\mu_S(t)S_tdt+\sigma_S(t)S_tdW_t\]
\[dD_t=\delta_S(t)S_tdt\]
\[dG_t=dS_t+dD_t\]
We must ensure $G^*$ is a $Q$-martingale to avoid arbitrage. Starting with the
dynamics of $dS_t^*$ under the $P$ probability measure. We do not need to do
this step for the dividend process because $D$ has no Weiner process.
Identification:
\[S_t^*=\frac{S_t}{A_t}=e^{-\int_t^Tr_udu}S_t=f(t,S)\]
\[
  dS_t^*=\left[ f_1+f_2+\frac{1}{2}f_{22}(\sigma_SS)^2\right] dt+\sigma_Sf_2dW_t
\]
Calculation:
\[
  f_1=e^{-\int_t^Tr_udu}\frac{\partial}{\partial t}
  \left( -\int_t^Tr_udu\right) S_t
\]
$S_t$ does not depend on $t$ it just has a subscript
\[=S_tA_t^{-1}\frac{\partial}{\partial t}-r_tdt=-r_tA_t^{-1}S_t\]
$A_t^{-1}$ does not depend on $S$
\[f_2=e^{-\int_t^Tr_udu}1=A_t^{-1}\]
$S$ dies in $f_2$ so there is nothing left to differentiate in $f_{22}$
\[f_{22}=0\]
Substitution:
\[
  dS_t^*=\left[ -r_tA_t^{-1}S_t+A_t^{-1}\mu_S(t)S_t+\frac{1}{2}0(\sigma_SS)^2
  \right] dt+A_t^{-1}\sigma_S(t)S_tdW_t
\]
Simplification: Remembering $S_t^*=\frac{S_t}{A_t}$
\[dS_t^*=\left[ \mu_S(t)S_t^*-r_tS_t^*\right] dt+\sigma_S(t)S_t^*dW_t\]
\[dS_t^*=\left[ \mu_S(t)-r_t\right] S_t^*dt+\sigma_S(t)S_t^*dW_t\]
This is only the dynamics of $S_t^*$ under $P$ probabilities with a standard
Weiner process.
We must make sure the discounted gains process $G^*$ is a $Q$-martingale.
\[dG_t^*=d(S_t^*+D_t^*)=dS_t^*+dD_t^*\]
\[=dS_t^*+d\int_0^t\frac{\delta_S(u)S_u}{A_u}du\]
Noticing $\frac{S_u}{A_u}=S_u^*$
\[=[\mu_S(t)-r_t]S_t^*dt+\sigma_S(t)S_t^*dW_t+\delta_S(t)S_t^*dt\]
Changing from $P$ to $Q$ probability measures with
$d\tilde{W}_t=dW_t+\lambda_tdt$
\[
  =[\mu_S(t)+\delta_S(t)-r_t]S_t^*dt+
  \sigma_S(t)S_t^*(d\tilde{W}_t-\lambda(t)dt)
\]
\[
  =[\mu_S(t)+\delta_S(t)-r_t-\sigma_S(t)\lambda(t)]S_t^*dt+
  \sigma_S(t)S_t^*d\tilde{W}_t
\]
In order to truly treat $d\tilde{W}_t$ as a standard Wiener and say we have
changed probabilities we need to choose $\lambda$ such that the drift of the
gains process $\mu_{G^*}=0$. Otherwise there would be arbitrage and much
wailing and gnashing of teeth.
\[\mu_S(t)+\delta_S(t)-r_t-\sigma_S(t)\lambda(t)=0\]
\[\lambda(t)=\frac{\mu_s(t)+\delta_S(t)-r_t}{\sigma_S(t)}\]
Where $\mu_S+\delta_S$ is the total return of the stock,
$\mu_S+\delta_S-r_t$ is the excess return and dividing the excess return by
$\sigma_S$ gives the Sharpe ratio.

We have identified $Q$ through its effect on $d\tilde{W}_t=dW_t+\lambda(t)$.
The payoff for an option, for example a call, is a function of $S_T$ and $K$,
not $S_t^*$ or any other process under $Q$ probabilities. We must determine
$S_T$ under $Q$ because
\begin{equation} \label{callExpVal}
  c_t=E_t^Q\left[ e^{-\int_t^Tr_udu}max(S_T-K) \right]
\end{equation}
We found the dynamics of our gains process under $Q$ in order to find $\lambda$
so that we can plug this into our stock price process once we change to $Q$
probabilities. We used the gains process to find $\lambda$ because all gains
processes must be $Q$-martingales to avoid arbitrage. Having found $\lambda$ we
can use it to adjust processes to make $d\tilde{W}_t$ a standard Wiener.
\[dS_t=\mu_S(t)S_tdt+\sigma_S(t)S_t[d\tilde{W}_t-\lambda(t)dt]\]
\[=[\mu_S(t)+\sigma_S(t)(-\lambda(t))]S_tdt+\sigma_S(t)S_td\tilde{W}_t\]
substituting for $\lambda$
\[
  \left[\mu_S(t)+\sigma_S(t)
  \left( -\frac{\mu_S(t)+\delta_S(t)-r_t}{\sigma_S(t)}\right)\right]
  S_tdt+\sigma_S(t)d\tilde{W}_t
\]
\[[r_t-\delta_S(t)]S_tdt+\sigma_S(t)S_td\tilde{W}_t\]
\textbf{No $\mu_S$ here} and we recognize it as a geometric Wiener process.
\[
  S_T=S_te^{\int_t^Tr(u)du-\delta_S(u)-\frac{1}{2}\sigma_S^2(u)du+
  \int_t^T\sigma_S(u)d\tilde{W}_u}
\]
and remembering \ref{callExpVal} we can use Feynman-Kac. We just change the
diffusion when using $Q$-probabilities. We have used the same replicating
portfolio logic to arrive here so we can use the same Black-Scholes-Merton
formula.
