
\section{Standard Weiner Process}

The standard Wiener process, $W$, is the whole stochastic process or law of
motion. Also referred to as Brownian motion after the biologist who first came
up with the informal idea while trying to describe the movements of pollen in
liquid. Wiener formalized the idea.
\begin{itemize}
    \item $W_t$: value at time t, unknown before t, often random variable
    \item $W_{\omega}$: particular path given state $\omega$
    \item $W_{t,\omega}$: value at time t, given state $\omega$
\end{itemize}
$W$ has the following properties;
\begin{enumerate}
\item $W_0=0$ (normalized to 0)
\item $W_t-W_s \sim N(0, t-s)$ (variance is equal to the time increment)
\item Non-overlapping time increments of $W$ are independent
\item $W$ has continuous sample paths, $W_{t, \omega}$ is continuous in t.
    This is possible because, while the normal distribution can jump to an
    infinite number of values, as the increments between jumps become smaller
    the variance converges to 0.
\end{enumerate}

Given the above properties, $W_{t,\omega}$ is not differentiable in t. This is
important because if we could find the derivative we would know which way the
process would move next and trade based on this knowledge. Not consistent with
no arbitrage. This opportunity would be traded away anyway.

We are not finished because W can be negative. This does not fit well with a
stock price process due to limited liability. Someone owning a stock can lose a
maximum of the amount they invested and no more. For a physical asset or
interest rate the possibility of a negative value is not a complete no go, but
we will focus on stocks.
